\sysfull

### 2018-08-02 paper/asplos19/intro.tex
New Makefile and more tiny intro figure

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\cite{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later---corresponds to {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
operation-power failure-restart cycles. As an example, recharge times may be
tens of seconds in radio frequency powered medical device~\cite[Fig.
3c]{nadeau_naturebio_2017}. The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.

\textbf{Data Consistency in Intermittent Computing.}  
%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device. In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}. In a failure period, 
Upon power failures, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM or Flash. While capturing periodic
checkpoints~\cite{mementos,quickrecall} and sleep
scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent execution to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\cite{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
expensive to design, verify, and manufacture. New architectures are also
inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today, but prior
software approaches have a number of limitations that are the focus of the work
in this paper. A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.

\begin{figure} %{t!}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation that should be executed despite arbitrarily-timed power failures.
%and these existing systems guarantee that each
%tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
%
The programmer explicitly expresses task-to-task control flow, or the compiler
converts existing program control-flow into task-to-task control flow to sequence 
the tasks that it defines.
%
Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

A savvy programmer may thus create very large tasks in an attempt to reduce
overhead by minimizing the number of task transitions. However, a very large
task may require more energy to complete than a device's fixed hardware energy
buffer can hold. Such a large task fails to completely execute assuming only
buffered energy is available (i.e., assuming the worst case). While some
environments may have sufficient harvestable energy to replenish a device's
capacitor during operation, prolonging the execution period and enabling the
task to complete, assuming cooperation from the physical environment to avoid
non-termination is a risky programming proposition. To eliminate this risk,
existing systems require the programmer or compiler to decompose a program
into small tasks, all of which complete using only buffered energy. These
constraints on task sizing leave an unsatisfying dilemma: large, efficient
tasks that risk non-termination, or small tasks that are guaranteed to
complete, but incur a high task transition and commit overhead. In this work,
we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads while avoiding the risk of non-termination.

@@

This paper introduces a new system called \sys, which executes a task-based
intermittent program by dynamically coalescing/splitting its tasks.
\sys accepts any static decomposition (i.e., from the programmer or from a compiler) 
and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute  
with no commit overhead between them, instead performing
task-end commit actions at the end of the second task only.
%
When there is sufficient energy to execute both tasks, two distinct, small
tasks are effectively executed as a single large task, preserving the atomicity
of both. If power fails during a coalesced task, execution restarts from the
{\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution. 
On the other hand, a large task that requires more energy to complete than a device buffered energy 
is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).

>>>>

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\cite{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later---corresponds to {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
operation-power failure-restart cycles. 
%As an example, recharge times may be tens of seconds in radio frequency powered medical device~\cite[Fig.3c]{nadeau_naturebio_2017}. 
The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.

\textbf{Data Consistency in Intermittent Computing.}
%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device. In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}. In a failure period,
Upon power failures, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM. While capturing periodic
checkpoints~\cite{mementos,quickrecall} and sleep
scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent execution to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\cite{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
expensive to design, verify, and manufacture. New architectures are also
inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today. Therefore, this work focuses on a key limitation of task-based software approaches, namely, the {\em inflexibility} of statically decomposing a program into tasks.

\begin{figure} %{t!}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-tiny.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation that should be executed despite arbitrarily-timed power failures.
%and these existing systems guarantee that each
%tasks executes {\em atomically}, despite arbitrarily-timed power failures.
%
The programmer (or a compiler) explicitly expresses task-to-task control flow.
%
Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

A programmer may thus create very large tasks in an attempt to reduce
overhead by minimizing the number of task transitions. However, a very large
task may require more energy to complete than a device's fixed hardware energy
buffer can hold which may lead to a program non-termination problem. To eliminate this risk,
existing systems require the programmer to decompose a program
into small tasks, all of which complete using only buffered energy. These
constraints on task sizing leave an unsatisfying dilemma: large, efficient
tasks that risk non-termination, or small tasks that are guaranteed to
complete, but incur a high task transition and commit overhead. In this work,
we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.

@@

This paper introduces a new system called \sys, which executes a task-based
intermittent program by dynamically coalescing/splitting its tasks.
\sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
with no commit overhead between them, instead performing
task-end commit actions at the end of the second task only.
%
When there is sufficient energy to execute both tasks, two distinct, small
tasks are effectively executed as a single large task, preserving the atomicity
of both. If power fails during a coalesced task, execution restarts from the
{\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
On the other hand, a large task that requires more energy to complete than a device buffered energy
is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
